This markdown will be used to run tests of different tidysdm models on the white shark data.

```{r}
suppressPackageStartupMessages(library(tidysdm))
for (f in list.files("/mnt/s1/projects/ecocast/projects/koliveira/subprojects/carcharodon/functions", pattern = "^.*\\.R$", full.names = TRUE)){
  source(f)
}
```

Block that loads all of my envi layers, observation data loaded. Then thin the background. Reinterpolate the envi data to one format.

We'll first pull our observations that we have from the gather_data.R script.
```{r}
bb = cofbb::get_bb("nefsc_carcharodon", form = "sf")
coast = rnaturalearth::ne_coastline(scale = "large", returnclass = "sf") |>
  sf::st_geometry()

vars = c("eventDate", "Year", "month", "basisOfRecord", "geom", "log_depth", "brick_tbtm", "brick_mld", "brick_sss", "brick_sbtm", "gseal", "hseal")

obs = read_brickman_points(file = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/get_data_workflow/versions/v01/v01.003/brickman_covar_obs_bg.gpkg") |>
  sf::st_as_sf() |>
  dplyr::filter(id == 1) |>
  dplyr::select(all_of(vars))
```

Next we'll pull the covariates. These are stored in the data directory.
```{r}
data_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/data"
depth_path = "brickman/bathy"
brick_path = "brickman/gom_carcharodon"
fish_path = "fish_data/MDAT_Fish_SummaryProducts_NEFSC/commondata/raster_data"
etopo_path = "mapping/etopo"

mon_no = seq(from = 1, to = 12)
depth = stars::read_stars(file.path(data_path, depth_path, "PRESENT_PRESENT_Bathy_depth_mon.tif")) |>
  dplyr::rename(log_depth = "PRESENT_PRESENT_Bathy_depth_mon.tif") |>
  dplyr::mutate(log_depth = log(log_depth))
dd = sapply(mon_no, function(mon) {depth}, simplify = FALSE)
depth = do.call(c, append(dd, list(along = list(band = mon_no)))) |>
  stars::st_set_dimensions("band", offset = NA, delta = NA)
tbtm = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_Tbtm_mon.tif")) |>
  dplyr::rename(tbtm = "PRESENT_PRESENT_Tbtm_mon.tif")
mld = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_MLD_mon.tif")) |>
  dplyr::rename(mld = "PRESENT_PRESENT_MLD_mon.tif")
sss = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_SSS_mon.tif")) |>
  dplyr::rename(sss = "PRESENT_PRESENT_SSS_mon.tif")
sbtm = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_Sbtm_mon.tif")) |>
  dplyr::rename(sbtm = "PRESENT_PRESENT_Sbtm_mon.tif")
gseal = load_seal(scenario = "PRESENT", year = NA, species = "gray") |>
  dplyr::rename(gseal = "prediction.tif")
hseal = load_seal(scenario = "PRESENT", species = "harbor") |>
  dplyr::rename(hseal = "prediction.tif")

dynamic_names = c("tbtm", "mld", "sss", "sbtm", "gseal", "hseal")
dynamic_preds = twinkle::bind_attrs(list(tbtm, mld, sss, sbtm, gseal, hseal)) |>
  set_names(dynamic_names)

preds = c(dynamic_preds |> units::drop_units() , depth |> units::drop_units(), along = NA_integer_, try_hard = TRUE)

list(preds = dynamic_preds, depth = depth) |> 
  saveRDS("bind_cluster.Rds") 

charlier::sendmail(to="btupper@bigelow.org", attachment = "bind_cluster.Rds")

preds = twinkle::bind_attrs(list(dynamic_preds, depth), along = "band")

dynamic_preds
static_preds
```

Set up a mask for our data
```{r}
mask = stars::read_stars(file.path(data_path, etopo_path, "etopo_warped_mask.tif")) |>
  rlang::set_names("mask") |>
  st_warp(dest = depth) |>
  mutate(mask = if_else(mask == 0, factor(NA_character_), factor(1))) 

plot(mask, breaks = "equal", axes = TRUE, reset = FALSE)
plot(sf::st_geometry(obs), pch = "+", add = TRUE)
plot(coast, col = "orange", add = TRUE)
```

Now we will thin our obs by using the function thin_by_cell(). 
```{r}
psat = obs |>
  filter(basisOfRecord == "PSAT")
non_psat_obs = obs |>
  filter(basisOfRecord != "PSAT")
thinned_psat = tidysdm::thin_by_cell(psat, raster = mask)
thinned_psat2 = tidysdm::thin_by_dist(thinned_psat, dist_min = km2m(20))
thinned_obs <- rbind(non_psat_obs, thinned_psat2)
plot(mask, breaks = "equal", axes = TRUE, reset = FALSE)
plot(sf::st_geometry(thinned_obs), pch = "+", add = TRUE)
plot(coast, col = "orange", add = TRUE)

```


```{r}
observation_density_raster = rasterize_point_density(obs, mask, dilate = 3)
pal = terra::map.pal("viridis", 50)
plot(observation_density_raster, 
     #col = pal, 
     #nbreaks = length(pal) + 1, 
     #breaks = "equal", 
     reset = FALSE)
plot(coast, col = "orange", add = TRUE)   
```

```{r}
model_input <- tidysdm::sample_background(
    data = thinned_obs, 
    raster = observation_density_raster,
    n = 2*nrow(thinned_obs),
    method = "bias",
    class_label = "background",
    return_pres = TRUE) |>
  dplyr::mutate(time = lubridate::NA_Date_, .after = 1)
ix = model_input$class == "presence"
model_input$time[ix] <- thinned_obs$eventDate

plot(model_input['class'], pch = 1, cex = 0.2,reset = FALSE, axes = TRUE)
plot(coast, col = "black", add = TRUE)
```

```{r}
nback = sum(!ix)
days_sample = sample_time(obs$eventDate, 
                          size = nback, 
                          by = "month", 
                          replace = TRUE, 
                          weighted = TRUE)
# recall ix is the logical identifying the class "presence"
model_input$time[!ix] <- days_sample
```


extract points data from covariates

issue coming from the way that preds is created, need to go back and modify that
```{r}
input_data = stars::st_extract(preds, at = model_input) |>
  sf::st_as_sf() |>
  dplyr::as_tibble() |>
  dplyr::select(dplyr::all_of(names(preds)))
model_input = dplyr::bind_cols(model_input, input_data) |>
  dplyr::select(-dplyr::all_of("time")) |>
  dplyr::relocate(dplyr::all_of("class"), .before = 1) |>
  dplyr::glimpse()
```


```{r}
na.omit(model_input) |>
  dplyr::select(dplyr::all_of(c("class", names(preds)))) |>
  tidysdm::plot_pres_vs_bg(class)
```

```{r}
model_input |>
  dplyr::select(dplyr::all_of(c("class", names(preds)))) |>
  na.omit() |>
  tidysdm::dist_pres_vs_bg(class)
```

```{r}
pairs(preds)
```

```{r}
vars_uncor <- tidysdm::filter_collinear(preds, cutoff = 0.7, method = "vif_cor", verbose = TRUE)
```

```{r}
vars_uncor
```

```{r}
vars_keep = c("mld", "hseal", "tbtm", "gseal", "sbtm")
model_input = dplyr::select(model_input,
                           dplyr::all_of(c("class", vars_keep))) |>
  na.omit()
model_input
#|>
#  sf::write_sf(file.path(data_path, "tidysdm/obs/model_input.gpkg"))
```

Setting up recipe
```{r}
rec = recipe(head(model_input), formula = class ~ .)
rec
```

```{r}
models = workflow_set(preproc = list(default = rec),
                      models = list(maxent = tidysdm::sdm_spec_maxent()),
                      cross = TRUE) |>
  option_add(control = tidysdm::control_ensemble_grid())
models
```

Cross folding
```{r}
input_cv = spatial_block_cv(data = model_input, v = 10, n = 5)
autoplot(input_cv)
```

```{r}
models = models |>
  workflow_map("tune_grid",
               resamples = input_cv,
               grid = 10,
               metrics = tidysdm::sdm_metric_set(),
               verbose = TRUE)
models
autoplot(models)
```
```{r}
ensemble = simple_ensemble() |>
  add_member(models, metric = "roc_auc")
ensemble
autoplot(ensemble)
```

```{r}
prediction_present = predict_raster(ensemble, slice(vars_keep, "time", 1))
plot(p, reset = FALSE, zlim = c(0,1), main = format(times[1], "%Y-%m-%d"), 
     breaks = seq(from = 0, to = 1, by = 0.1), 
     axes = TRUE)
plot(coast, add = TRUE, col = "orange")

```




```{r}
predict_by_mon(obs = obs, 
               dyn = dynamic_preds, 
               static = static_preds, 
               mask = mask, 
               figure_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/tidysdm_workflow/figures",
               tidy_data_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/data/tidysdm")

if(FALSE){
  mon = 8
  obs = obs
  dyn = dynamic_preds
  static = static_preds 
  mask = mask
  figure_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/tidysdm_workflow/figures"
  tidy_data_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/data/tidysdm"
}
```
