This markdown will be used to run tests of different tidysdm models on the white shark data.

```{r}
suppressPackageStartupMessages(library(tidysdm))
for (f in list.files("/mnt/s1/projects/ecocast/projects/koliveira/subprojects/carcharodon/functions", pattern = "^.*\\.R$", full.names = TRUE)){
  source(f)
}
```

Block that loads all of my envi layers, observation data loaded. Then thin the background. Reinterpolate the envi data to one format.

We'll first pull our observations that we have from the gather_data.R script.
```{r}
bb = cofbb::get_bb("nefsc_carcharodon", form = "sf")
coast = rnaturalearth::ne_coastline(scale = "large", returnclass = "sf") |>
  sf::st_geometry()

vars = c("eventDate", "Year", "month", "basisOfRecord", "geom", "log_depth", "brick_tbtm", "brick_mld", "brick_sss", "brick_sbtm", "gseal", "hseal")

obs = read_brickman_points(file = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/get_data_workflow/versions/v01/v01.003/brickman_covar_obs_bg.gpkg") |>
  sf::st_as_sf() |>
  dplyr::filter(id == 1) |>
  dplyr::select(all_of(vars))
```

Next we'll pull the covariates. These are stored in the data directory.
```{r}
data_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/data"
depth_path = "brickman/bathy"
brick_path = "brickman/gom_carcharodon"
fish_path = "fish_data/MDAT_Fish_SummaryProducts_NEFSC/commondata/raster_data"
etopo_path = "mapping/etopo"

mon_no = seq(from = 1, to = 12)
depth = stars::read_stars(file.path(data_path, depth_path, "PRESENT_PRESENT_Bathy_depth_mon.tif")) |>
  dplyr::rename(log_depth = "PRESENT_PRESENT_Bathy_depth_mon.tif") |>
  dplyr::mutate(log_depth = log(log_depth))
dd = sapply(mon_no, function(mon) {depth}, simplify = FALSE)
depth = do.call(c, append(dd, list(along = list(band = mon_no)))) |>
  stars::st_set_dimensions("band", offset = NA_real_, delta = NA_real_)
tbtm = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_Tbtm_mon.tif")) |>
  dplyr::rename(tbtm = "PRESENT_PRESENT_Tbtm_mon.tif")
mld = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_MLD_mon.tif")) |>
  dplyr::rename(mld = "PRESENT_PRESENT_MLD_mon.tif")
sss = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_SSS_mon.tif")) |>
  dplyr::rename(sss = "PRESENT_PRESENT_SSS_mon.tif")
sbtm = stars::read_stars(file.path(data_path, brick_path, "PRESENT_PRESENT_Sbtm_mon.tif")) |>
  dplyr::rename(sbtm = "PRESENT_PRESENT_Sbtm_mon.tif")
gseal = load_seal(scenario = "PRESENT", year = NA, species = "gray") |>
  dplyr::rename(gseal = "prediction.tif")
hseal = load_seal(scenario = "PRESENT", species = "harbor") |>
  dplyr::rename(hseal = "prediction.tif")

dynamic_names = c("tbtm", "mld", "sss", "sbtm", "gseal", "hseal")
dynamic_preds = twinkle::bind_attrs(list(tbtm, mld, sss, sbtm, gseal, hseal)) |>
  set_names(dynamic_names)

preds = c(dynamic_preds,  depth)

```

Set up a mask for our data
```{r}
mask = stars::read_stars(file.path(data_path, etopo_path, "etopo_warped_mask.tif")) |>
  rlang::set_names("mask") |>
  stars::st_warp(dest = dplyr::slice(depth, band, 1)) |>
  dplyr::mutate(mask = if_else(mask == 0, factor(NA_character_), factor(1))) 

plot(mask, breaks = "equal", axes = TRUE, reset = FALSE)
plot(sf::st_geometry(obs), pch = "+", add = TRUE)
plot(coast, col = "orange", add = TRUE)
```

Now we will thin our obs by using the function thin_by_cell(). 
```{r}
psat = obs |>
  dplyr::filter(basisOfRecord == "PSAT")
non_psat_obs = obs |>
  dplyr::filter(basisOfRecord != "PSAT")
thinned_psat = tidysdm::thin_by_cell(psat, raster = mask)
thinned_psat2 = tidysdm::thin_by_dist(thinned_psat, dist_min = km2m(20))
thinned_obs <- rbind(non_psat_obs, thinned_psat2)
plot(mask, breaks = "equal", axes = TRUE, reset = FALSE)
plot(sf::st_geometry(thinned_obs), pch = "+", add = TRUE)
plot(coast, col = "orange", add = TRUE)

```

```{r}
predict_by_mon(obs = obs, 
               thinned = thinned_obs, 
               preds = preds, 
               mask = mask, 
               figure_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/tidysdm_workflow/figures",
               tidy_data_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/data/tidysdm",
               model_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/tidysdm_workflow/models",
               v = 10)

```

```{r}
model_path = "/mnt/ecocast/projects/koliveira/subprojects/carcharodon/workflows/tidysdm_workflow/models"
aug_brt_model = read_model(filename = file.path(model_path, "8/gbm_model.rds"))
aug_maxent_model = read_model(filename = file.path(model_path, "8/maxent_model.rds"))
aug_rf_model = read_model(filename = file.path(model_path, "8/rf_model.rds"))
aug_tss_ensemble = read_ensemble(filename = file.path(model_path, "8/tss_ensemble.rds"))
vars_keep = c("mld", "hseal", "tbtm", "gseal", "sbtm", "log_depth")
aug_preds = dplyr::select(preds, dplyr::all_of(vars_keep))
aug_tss_predict_test = predict_raster(z, aug_preds, probability = TRUE)
plot(aug_tss_predict_test, reset = FALSE, 
     breaks = seq(from = 0, to = 1, by = 0.1))
plot(coast, col = "orange", add = TRUE)

```





